---
title: "Lab-3.Data-Wrangling.Rmd"
author: "Liaqat Ali"
date: "9/14/2020"
output: html_document
---


```{r}
library(tidyverse)
library(dplyr)
```

```{r, eval= FALSE}
 download.file(url = "https://ndownloader.figshare.com/files/2292169",
              destfile = "data/portal_data_joined.csv")

```

```{r}
surveys <- read.csv("data/portal_data_joined.csv")
```

```{r}
head(surveys)
```


```{r, eval=FALSE}
View(surveys)
```


```{r}
head(surveys)
```

```{r}
str(surveys)
```

```{r}
nrow(surveys)
ncol(surveys)
```

```{r}
head(surveys)
tail(surveys)
```

```{r}
names(surveys)
```

```{r}
summary(surveys)
```
# Challenge


Inspecting data.frame Objects

###  What is the class of the object surveys?
##   How many rows and how many columns are in this object?
#    How many species have been recorded during these surveys?

```{r}
str(surveys)
```

## * class: data frame
## * how many rows: 34786,  how many columns: 13
## * how many species: 48


Indexing and subsetting data frames

```{r}
# first element in the first column of the data frame (as a vector)
surveys[1, 1]   
# first element in the 6th column (as a vector)
surveys[1, 6]   
# first column of the data frame (as a vector)
surveys[3,2 ]    
# first three elements in the 7th column (as a vector)
surveys[1:3, 7] 
# the 3rd row of the data frame (as a data.frame)
surveys[3, ]    
# equivalent to head_surveys <- head(surveys)
head_surveys <- surveys[1:6, ] 
```

```{r}
surveys[10,10] # the whole data frme, except the first column
```

```{r}
surveys[-(7:34786), ] # Equivalent to head(surveys)
```

surveys["species_id"]
surveys[, "species_id"]
surveys[["species_id"]]
surveys$species_id

# Chanllenge

```{r}
## 1.
surveys_200 <- surveys[200, ]
## 2.
# Saving `n_rows` to improve readability and reduce duplication
n_rows <- nrow(surveys)
surveys_last <- surveys[n_rows, ]
## 3.
surveys_middle <- surveys[n_rows / 2, ]
## 4.
surveys_head <- surveys[-(7:n_rows), ]
```

## Factors
```{r}
sex <- factor (c("male", "female", "female", "male"))
levels(sex)
nlevels(sex)
```


sex # current order

#> [1] male female female male
#> Levels: female male

sex <- factor (sex, levels = c("male", "female"))
sex # after re-ordering

#> [1] male female female male
#. Levels: male female

year_fct <- factor(c(1990, 1983, 1977, 1998, 1990))
as.numeric(year_fct)
as.numeric(as.character(year_fct))
as.numeric(levels(years_fct))[year_fct]


```{r}
plot(as.factor(surveys$sex))
```



```{r}
sex <- factor(surveys$sex)
head(sex)

#> [1] M M
#> Levels: F M

levels(sex)

#> [1] "" "F" "M"

levels(sex)[1] <- "undetermined"
levels(sex)

#> [1] "undetermined" "F"       "M"
head(sex)

#> [1] M      M        undetermined undetermined undetermined
#> [6] undetermined
#> Levels: undetermined F M
```

## Challenge

```{r }
levels(sex)[2:3] <- c("female", "male")
sex <- factor(sex, levels = c("female", "male", "undetermined"))
plot(sex)
```


```{r }
   animal_data <- data.frame(
   animal = c("dog", "cat", "sea cucumber", "sea urchin"),
   feel = c("furry","furry", "squishy", "spiny"),
   weight = c(45, 8, 1.1, 0.8)
   )
   
```

```{r }
## country_climate <- data.frame(country=c("Canada","Panama", "South Africa", "Australia"),
       ## climate = c("cold", "hot", "temperate", "hot/temperate"),
       ## temperature = c(10, 30, 18, "15"),
       ## northern_hemisphere = c(TRUE, TRUE, FALSE, "FALSE"),
       ## has_kangaroo = c(FALSE, FALSE, FALSE, 1)
  ```


```{r }
library(tidyverse)
library(lubridate)
```

```{r}
summary(surveys$date)
```



